Temperature and wave exposure as drivers of morphology in two kelp species

### 2019-08-04 morphology_paper.Rmd
new sections added in results for redundancy analysis and html output updated

---
title: Temperature and wave exposure as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 7
    fig_width: 7
  word_document:
    reference_docx: morphology_paper_files/frontiers_template.docx
  html_notebook:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
papersize: A4
tables: yes
fontsize: 10pt
---

@@

```{r echo=FALSE, fig.cap='Schematic of *E. maxima* on the left and *L. pallida* on the right. (Dyer 2018)'}
library(png)
library(grid)

@@

# The patched Roman Rock temperatures
load("data/rr_patch.Rdata")

@@

Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St. Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay ([Fig. 4]. Annual peak wave period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak wave period amongst all sites. These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.  Annual maximum peak wave period ranged from 18.9 s at St. Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula. Data for median wave direction showed no trend across timescales; however the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter. Summer data were the exception and showed lower variation for the eastern side of the peninsula [Fig. 4]. Wave variables divided into Annual (top), Austral summer (middle) and Austral winter (bottom) respectively. This phenomenon can be explained by the sheltering effect of the Cape Peninsula, resulting in a wave climate with a narrower directional spreading on the eastern side of the peninsula (refer to Fig. 7).

```{r wave power map, eval=FALSE, fig.cap='wave power map', include=FALSE}
library(png)
library(grid)

img_wavepower <- readPNG("figures/CapePen_Total.PNG")
legend_wave_power <- readPNG("figures/CapePen_Legend.PNG")
grid.raster(img_wavepower,legend_wave_power)
```
**insert wave exposure map here**

In Figure 5 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length). The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in [Fig. 6]. The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits. The western periphery of the Cape peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures [Fig. 6]. THere the marked seasonality, with higher energy waves during winter, may be clearly observed once more. To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in [Fig. 7]. Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered. It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions. This results in positive and negative wave exposure anomalies all around the peninsula.

@@

```{r rda analysis lamanaria, eval=TRUE, include=FALSE, message=FALSE, warning=FALSE}
#data processed in the temperature and wave sections will be used here eg. abio_clim
# remove unnecessary columns
lam_df <- subset(lam_data, select = -c(1,2,6))

@@

```{r rda deep ecklonia processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

@@

@@

Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of *L. pallida* morphological characteristics with an adjusted *R*2 of 0.73, explaining 85% of the variation (global permutation test on final model: d.f = 4, *F* = 7.18, *p* = 0.001). The model consisted of two significant axes: RDA1 and RDA2 explained 66% and 24% of the variation, respectively. bla bla

Forward selection, assessment of VIF and pearson correlation coefficients retained *E. maxima* morphological characteristics with an adjusted *R*2 of 0.77, explaining 92% of the variation (global permutation test on final model: d.f = 8, *F* = 6.08, *p* = 0.001). bla bla

>>>>

---
title: Temperature and wave exposure as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
    fig_crop: no
    fig_height: 7
    fig_width: 7
  word_document:
    reference_docx: frontiers_template.docx
  html_notebook:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
papersize: A4
tables: yes
fontsize: 10pt
---

@@

```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.'}
library(png)
library(grid)

@@

# The patched Roman Rock temperatures
load("data/temp_data/rr_patch.Rdata")

@@

```{r wave rose figure, echo=FALSE, eval= FALSE, fig.cap='Wave rose figure'}

img_wave_rose <- readPNG("figures/wave_rose.png")
grid.raster(wave rose)
```
**insert wave roses here**
Annual mean wave peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St. Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay ([Fig. 4]. Annual peak wave period was 10.0 ± 3.0 s and 9.2 ± 3.1 s in winter and summer respectively, Miller’s point and had the lowest peak wave period amongst all sites. These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side of the peninsula, with Miller’s point showing the highest variation across timescales.  Annual maximum peak wave period ranged from 18.9 s at St. Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula. Data for median wave direction showed no trend across timescales; however the variation (SD) in wave direction is lower on the western side of the peninsula compared to the eastern side both annually and for winter. Summer data were the exception and showed lower variation for the eastern side of the peninsula [Fig. 4]. Wave variables divided into Annual (top), Austral summer (middle) and Austral winter (bottom) respectively. This phenomenon can be explained by the sheltering effect of the Cape Peninsula, resulting in a wave climate with a narrower directional spreading on the eastern side of the peninsula (refer to Fig. 7).

```{r wave power map, echo=FALSE, eval= FALSE, fig.cap='Wave power map'}

library(png)
library(grid)

img_wavepower <- readPNG("figures/CapePen_Total.PNG")
legend_wave_power <- readPNG("figures/CapePen_Legend.PNG")
grid.raster(img_wavepower,legend_wave_power)
```
**insert wave power map here**
In Figure 5 the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length). The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in [Fig. 6]. The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits. The western periphery of the Cape peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures [Fig. 6]. THere the marked seasonality, with higher energy waves during winter, may be clearly observed once more. To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in [Fig. 7]. Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered. It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions. This results in positive and negative wave exposure anomalies all around the peninsula.

@@

```{r rda analysis lamanaria, message=FALSE, warning=FALSE, include=FALSE}
#data processed in the temperature and wave sections will be used here eg. abio_clim
# remove unnecessary columns
lam_df <- subset(lam_data, select = -c(1,2,6))

@@

Forward selection, assessment of VIF and an examination of pairwise Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of *L. pallida* morphological characteristics with an adjusted *R*2 of 0.49, explaining 72% of the variation (global permutation test on final model: d.f = 1, *F* = 8.51, *p* = 0.008). RDA1 was the only significant axis in the model and explained 48% of the variation while RDA2 was not significant and explained 63% of the variation. Total length, stipe length, stipe diameter and stipe mass were postivley influenced (i.e., increased size corresponding with the environmental driver) by *hs_mean* and *dir_median* and negativley influenced (i.e., decreased size corresponding with the environmental driver) by *tp_sd* and *mean_temp*. Study sites from the eastern region clustred close together while no clear clustering was evident for the western region.

```{r rda deep ecklonia processing, message=FALSE, warning=FALSE, include=FALSE}

@@

Forward selection, assessment of VIF and Pearson correlation coefficients allowed us to retained as the most parsimonious descriptors of deep *E. maxima* morphological characteristics with an adjusted *R*2 of 0.39, explaining 67% of the variation (global permutation test on final model: d.f = 5, *F* = 2.4, *p* = 0.026). RDA1 was the only significant axis in the model (p = 0.03) and explained 72% of the variation. The non-significant axis for the model explained 17% of the variation. Primary length, primary width and stipe circumference were positivley influenced by *hs_mean*, *tp_mean* and *dir_median* for for kelps from the western region. The remaining morphological characteristics were postivley inlfuenced by *min_temp* and negatively influenced by *tp_sd* for kelps ffrom the eastern region.

@@
